Verifying Spatial Properties of Array Computations

### 2017-04-13 oopsla17/lattice-model.tex
semantic model for specs

Third, we need to mark presence of modifiers such as \texttt{readOnce} and
\texttt{atMost} as introduced in \Cref{}. In the context of
multiplicity modifiers we have the following semantics:
%
\begin{align*}
  \interp{\texttt{readOnce}} & = \mathsf{mult} &
  \llbracket\epsilon\rrbracket & = \mathsf{once}
\end{align*}

In the context of approximation modifiers, we have
%
\begin{align*}
  \interp{\texttt{atLeast}} & = \mathsf{lower} &
  \interp{\texttt{atMost}} & = \mathsf{upper} &
  \llbracket \epsilon \rrbracket = \mathsf{exact}
\end{align*}

Finally, we can give full semantics of a stencil specification:
%
\begin{equation*}
  \interp{\texttt{stencil \textcap{Mult}, \textcap{Appr}, \textcap{Region}}}_N =
    \interp{\textcap{Mult}}
           {(\interp{\textcap{Appr}}
                    {(\interp{\textcap{Region}}_N)})}
\end{equation*}

>>>>

We mark the presence of modifiers such as \texttt{readOnce} and
\texttt{atMost} as introduced in \Cref{}. Multiplicity
modifiers are interpreted as injections in the $\mathsf{Mult}$
variant by $\interpMult{}$ in Fig~\ref{fig:semantics}.
In a similar way, approximation modifiers are interpreted
as injections into the $\mathsf{Approx}$ variant by $\interpApprox{}$.

Finally, we can give full semantics of a stencil specification:
%
\begin{equation*}
  \interp{\texttt{stencil \textcap{mult}, \textcap{approx}, \textcap{region}}}_N =
    \interpMult{\textcap{mult}}
           {(\interpMax{\textcap{approx}}
                    {(\interp{\textcap{region}}_N)})}
\end{equation*}

### 2017-04-13 oopsla17/oopsla17.tex
semantic model for specs

\noindent
An interpretation function $\interp{-}_N$ provides a semantics to
specifications as a member of $\textsf{Mult} (\textsf{Approx} (\mathcal{P}{\bz}^N)))$
(where $n$ is the maximum dimensionality of the specification being
modelled) mapping closed\footnote{That is, we assume there are no
  occurrences of \textit{rvar} in a specification being modelled.  Any
  \emph{open} specification containing region variables can be made
  closed by straightforward syntactic substitution with a (closed)
  \textit{region}.} specifications to sets of
$n$-dimensional index schemes. The interpretation is overloaded on
\emph{regions} in \Cref{fig:region-model} and on the top-level
of a specification \textit{spec} in \Cref{fig:spatial-model}.
Various intermediate notions are used.

>>>>

\noindent
An interpretation function $\interp{-}_N$ maps closed\footnote{That
  is, we assume there are no occurrences of \textit{rvar} in a
  specification being modelled.  Any \emph{open} specification
  containing region variables can be made closed by straightforward
  syntactic substitution with a (closed) \textit{region}.}
specifications to sets of $N$-dimensional index schemes with modifier
information, \ie{} specifications are mapped to
$\textsf{Mult} (\textsf{Approx} (\mathcal{P}({\bz})^N))$ where $N$ is
the maximum dimensionality of the specification being modelled.

The interpretation is overloaded on \emph{regions} in
\Cref{subfig:region-model}. Various intermediate notions are used.

\begin{defn}
Let $\textit{promote}_N : \mathbb{N}^+ \times \textit{Interval} \to
\textit{Interval}^N$ be a function generating an $N$-dimensional 
interval such that $\pi_i(\vecgen{N}{i}{\interv{a}{b}{c}}) = 
\interv{a}{b}{c}$ and $\pi_j(\vecgen{N}{i}{\interv{a}{b}{c}}) = 
\bz{}$ in all other dimensions $j$. 
\end{defn}

%
\begin{figure}[!t]
\begin{subfigure}[t]{0.5\textwidth}
\begin{align*}
\interp{-}_N & : \textit{region} \rightarrow \mathcal{P}(\bz)^N \\
  \interp{\stencil{p}{$i$}{}{}}_N & =
    \vecgen{N}{i}{\interv{0}{0}{\mathsf{true}}}\\
%
  \interp{\stencil{c}{$i$}{$k$}{\textcap{p}}}_N & =
    \vecgen{N}{i}{\interv{-k}{k}{\interp{\textcap{p}}}} \\
%
  \interp{\stencil{f}{$i$}{$k$}{\textcap{p}}}_N & =
    \vecgen{N}{i}{\interv{0}{k}{\interp{\textcap{p}}}} \\
%
  \interp{\stencil{b}{$i$}{$k$}{\textcap{p}}}_N & =
  \vecgen{N}{i}{\interv{-k}{0}{\interp{\textcap{p}}}}
\\
  \interp{\texttt{\textcap{r} + \textcap{s}}}_N & =
    \interp{\textcap{r}}_N \vee \interp{\textcap{s}}_N
\\
  \interp{\texttt{\textcap{r} * \textcap{s}}}_N & =
    \interp{\textcap{r}}_N \wedge \interp{\textcap{s}}_N \\[-1em]
\end{align*}
\caption{Interpretation of regions}
\label{subfig:region-model}
\end{subfigure}
\hspace{1em}
\begin{subfigure}[t]{0.4\textwidth}
\begin{align*}
\interpApprox{-} & : \textit{approx} \rightarrow (A \rightarrow
  \textsf{Approx} \, A) \\
\interpApprox{\texttt{atLeast}} & = \mathsf{lower} \\
  \interpApprox{\texttt{atMost}} & = \mathsf{upper} \\
  \interpApprox{\epsilon} & = \mathsf{exact} \\ \\
  \interpMult{-} & : \textit{mult} \rightarrow (A \rightarrow
  \textsf{Mult} \, A) \\
  \interpMult{\texttt{readOnce}} & = \mathsf{once} \\
  \interpMult{\epsilon} & = \mathsf{mult}
\end{align*}
\caption{Interpretation of modifiers}
\label{subfig:modifier-model}
\end{subfigure}
\label{fig:semantics}
\caption{Semantic model of specifications}
\end{figure}

The first four equations of \Cref{subfig:region-model} model region 
contants. The final two equations model the \term{+} and \term{*}
operators in terms of the join (union) and meet (intersection)
of interval schemes. Thus regions are modelled as members
of $\mathcal{P}(\bz)^N$.

We mark in our model the presence of modifiers such as
\texttt{readOnce} and \texttt{atMost} as introduced in \Cref{}.
Approximation modifiers are interpreted as injections into the
$\mathsf{Approx}$ variant by $\interpApprox{}$ in
\Cref{subfig:modifier-model}.  The $\textsf{Approx}$ type corresponds
to the presence or absence of the spatial approximation modifier, with
\textsf{exact} when there is no such modifier and \textsf{lower} and
\textsf{upper} for \term{atLeast} and \term{atMost}. In a similar way,
multiplicity modifiers are interpreted as injections in the
$\mathsf{Mult}$ variant by $\interpMult{}$, corresponding to the to
the presence or absence of the \term{readOnce} modifier as shown in
\Cref{subfig:modifier-model}.

\begin{defn}[Semantics of specifications]
The intermediate interpretations of \Cref{fig:semantics}
are composed to give a model for the top-level specification
syntax as:
%
\begin{equation*}
  \interp{\texttt{stencil \textcap{mult}, \textcap{approx}, \textcap{region}}}_N =
    \interpMult{\textcap{mult}} \;
           {(\interpApprox{\textcap{approx}} \;
                    {\interp{\textcap{region}}_N)}}
\end{equation*}
\end{defn}
%
\begin{thm}
The semantic model $\interp{-}_N$ of $N$-dimensional specifications
is sound with respects to the equational theory of the language,
that is:
%
\begin{equation*}
\forall S, T, N . \quad
S \equiv T \; \Rightarrow \;
\interp{S}_N = \interp{T}_N
\end{equation*}
%
We define the equational theory in \Cref{sec:equational-theory}.
\end{thm}
